---
title: "usdatasets: A Comprehensive Collection of U.S. Datasets"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{usdatasets: A Comprehensive Collection of U.S. Datasets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 10,  
  fig.height = 6 
)
```

```{r setup,include = FALSE}
library(usdatasets)

library(ggplot2)

library(dplyr)
```


# Introduction

The `usdatasets` package provides a **comprehensive collection of U.S. datasets, encompassing various fields such as crime, economics, education, finance, energy, healthcare, and more.** This package serves as a valuable resource for researchers and analysts seeking to perform in-depth analyses and derive insights from U.S.-specific data.

# Dataset Suffixes

To facilitate the identification of data types, a suffix is added to the end of the name of each dataset. These suffixes indicate the format and type of the datasets, such as:

- `tbl_df`: A tibble data frame
- `df`: A standard data frame
- `ts`: A time series object
- `matrix`: A matrix object
- `character`: A character vector
- `numeric`: A numeric vector
- `factor`: A factor variable

# Example Datasets

Here are some examples of datasets included in the `usdatasets` package:

- **marathon_tbl_df**: A tibble containing marathon race data, including runner statistics and performance metrics.
  
- **mn_police_use_of_force_df**: A data frame documenting incidents of police use of force in Minnesota.

- **nba_players_19_tbl_df**: A tibble that includes data on NBA players for the 2019 season.

- **ncbirths_tbl_df**: A tibble summarizing birth statistics across various demographics.

- **nyc_marathon_tbl_df**: A tibble containing results and statistics from the New York City Marathon.

- **nycvehiclethefts_tbl_df**: A data frame documenting vehicle theft incidents in New York City.


# Visualizing Data with ggplot2

To illustrate the data, we can use the `ggplot2` package to create some visualizations. Here are a few examples:

## 1. Visualization of Marathon Finish Times

```{r, message=FALSE, warning=FALSE}

# Example: Visualizing finish times of the NYC Marathon
# Ajustado para las columnas disponibles en 'marathon_tbl_df'
marathon_tbl_df %>%
  ggplot(aes(x = year, y = time, color = gender)) +
  geom_point(alpha = 0.6) +
  labs(title = "Marathon Finish Times by Year and Gender",
       x = "Year",
       y = "Finish Time (minutes)",
       color = "Gender") +
  theme_minimal()

```

## 2. Visualization of NBA Player Heights

```{r, message=FALSE, warning=FALSE}

# Example: Visualizing the distribution of NBA player heights
nba_players_19_tbl_df %>%
  ggplot(aes(x = height)) +
  geom_histogram(binwidth = 2, alpha = 0.7, fill = "blue", color = "black") +
  labs(title = "Distribution of NBA Player Heights",
       x = "Height (inches)",
       y = "Count") +
  theme_minimal()


```

## 3. Visualization of Police Use of Force Incidents

```{r, message=FALSE, warning=FALSE}

# Example: Visualizing police use of force incidents by race
mn_police_use_of_force_df %>%
  group_by(race) %>%
  summarize(count = n()) %>%
  ggplot(aes(x = reorder(race, count), y = count, fill = race)) +
  geom_bar(stat = "identity") +
  labs(title = "Incidents of Police Use of Force by Race",
       x = "Race",
       y = "Number of Incidents") +
  theme_minimal() +
  coord_flip()


```

## Conclusion

The `usdatasets` package is an invaluable tool for those looking to analyze and derive insights from a variety of U.S.-specific datasets. The suffixes used in the dataset names help users quickly identify the type of data they are working with, facilitating a smoother analysis process.

For more information and to explore the datasets, please refer to the package documentation.

<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>
